---
title: "exp1-analysis"
output: html_document
date: "2026-02-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(file = "./utils/exp1-utils.R") # loads all necessary packages
source(file = "./utils/exp1-deltaS-utils.R")
source(file = "./utils/stimuli.R")
```

This file has initial computation including variables like alpha for variance fit.

# Malagasy dataset
### Analysis of scaled data collection
n = 50

```{r data-load}
dataPath <- "data/exp1/csv/"
mgdem <- read_csv(paste(dataPath, "mg-dem.csv", sep = ""), show_col_types = FALSE)
mgdf <- read_csv(paste(dataPath, "mg-df.csv", sep = ""), show_col_types = FALSE)
mgdf_w <- read_csv(paste(dataPath, "mg-df-summary.csv", sep = ""), show_col_types = FALSE)

usdem <- read_csv(paste(dataPath, "us-dem.csv", sep = ""), show_col_types = FALSE)
usdf <- read_csv(paste(dataPath, "us-df.csv", sep = ""), show_col_types = FALSE)
usdf_w <- read_csv(paste(dataPath, "us-df-summary.csv", sep = ""), show_col_types = FALSE)

rm(dataPath)
```

### Analyzing within group
```{r plot assoc. weights}
## Association weight MDG
mgdf_w$con <- factor(mgdf_w$concept, levels = conceptListEn3) # to order the facet strips
p1 <- plotWeight_Err(mgdf_w, "Mean weight of assoc, lib: BCP37, MDG")
p1

## Association weight USA
usdf_w$con <- factor(usdf_w$concept, levels = conceptListEn3) # to order the facet strips
p2 <- plotWeight_Err(usdf_w, "Mean weight of assoc., lib: BCP37, USA") +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank())
p2

p12 <- grid.arrange(p1, p2, nrow = 1)
p12
```

```{r eval=FALSE}
# ggsave(file="./fig/weight-mg.png", plot = p1, width=10, height=16)
# ggsave(file="./fig/weight-us.png", plot = p2, width=10, height=16)
# ggsave(file="./output/fig/weight-both-unsorted.png", plot = p12, width=16, height=18)
```

#### Unique pairs of concepts and pairs of colors without repetition 
{red, yellow} == {yellow, red}
```{r}
concept_pairs <- as.data.frame(t(combn(unique(usdf$concept), 2))) %>% rename(concept1 = V1, concept2 = V2)
color_pairs <- as.data.frame(t(combn(unique(mgdf$color), 2))) %>% rename(color1 = V1, color2 = V2)
```

## Alphas
Estimating the value of alphas in the fit for standard deviation.

```{r US data}
# Aggregated data (x = average rating for each color of each concept)
usdf_agg <- usdf %>% group_by(concept, color) %>%
  summarise(xbar = mean(rating), sd_sample = sd(rating), .groups = "drop")

a_us <- round(coef(getFitSd(usdf_agg)), digits = 2) # alpha a_us = 1.384426 which is ~ 1.4
plotFitSd(a_us, usdf_agg, "USA data")
curve(a_both * x * (1 - x), add = TRUE, col = "red", lwd = 2)

# a_us2 <- round(coef(getFitSd2(usdf_agg)), digits = 2) # alpha a = 0.38, beta b = 0.32
# plotFitSd2(a_us2, usdf_agg, "USA")
# curve(a_us * x * (1 - x), add = TRUE, col = "orange", lwd = 2)

# -> The first model w/ a_us (s = 1.4 * xbar * (1-xbar)) fits the data better, so we will use a_us from here
# Note that in the paper, a_us is set to 1.4. Model is s = 1.4 * xbar * (1-xbar)

# xUs <- getXvalues(usdf_agg, concept_pairs, color_pairs, a_us) %>% mutate(deltaX = (xbar1 + xbar4) - (xbar2 + xbar3))
# get deltaX values with optimal alpha
xUs <- getXvalues_opt(usdf_agg, concept_pairs, color_pairs, 1.4) %>% 
  mutate(deltaX = (xbar1 + xbar4) - (xbar2 + xbar3)) # %>% 
  mutate(sd_diff1 = sd_sample1 - sd_fit1)

### Verifying if the outputs of getXvalues and getXvalues_opt are the same
# xus <- xUs[order(names(xUs))]; xus <- xus[order(row.names(xus)), ]
# xus2 <- xUs2[order(names(xUs2))]; xus2 <- xus2[order(row.names(xus2)),]
# anti_join(xus, xus2) # if nrow here == 0 then all rows in xus can be found in xus2, i.e. the two df are identical
# nrow(merge(xus, xus2))
### -> Yes, they are the same

deltaS_us <- getDeltaS(xUs) %>% mutate(deltaSdiff = deltaS_fit - deltaS_sample) 
summary(deltaS_us$deltaSdiff); sd(deltaS_us$deltaSdiff)

deltaS_us <- deltaS_us %>% select(concept_a, concept_b, color1, color2, deltaS_sample, deltaS_fit) 
# mean(deltaS_us$deltaS_sample - deltaS_us$deltaS_fit) ~= -0.007, i.e. the fitted S is correct
```


```{r MG data}
mgdf_agg <- mgdf %>% group_by(concept, color) %>%
  summarise(xbar = mean(rating), sd_sample = sd(rating), .groups = "drop")

a_mg <- round(coef(getFitSd(mgdf_agg)), digits = 2) # a = 1.536
plotFitSd(a_mg, mgdf_agg, "MDG data")
curve(1.4 * x * (1 - x), add = TRUE, col = "green", lwd = 2)

# a_mg2 <- round(coef(getFitSd2(mgdf_agg)), digits = 2)
# plotFitSd2(a_mg2, mgdf_agg, "MDG data")
# curve(a_mg * x * (1 - x), add = TRUE, col = "orange", lwd = 2)

# The 1st model w/ a_mg fits the data better, so we'll use a_mg from here
# sd_diff_mg <- a_mg*mgdf_agg$xbar*(1-mgdf_agg$xbar) - mgdf_agg$sd_sample; mean(sd_diff_mg)
# i=9; a_mg*mgdf_agg$xbar[i]*(1-mgdf_agg$xbar[i]) - mgdf_agg$sd_sample[i]

# get deltaX values with optimal alpha
xMg <- getXvalues_opt(mgdf_agg, concept_pairs, color_pairs, a_mg) %>% mutate(deltaX = (xbar1 + xbar4) - (xbar2 + xbar3)) #%>% 
  mutate(sd_diff1 = sd_sample1 - sd_fit1)
### Verifying if the outputs of getXvalues and getXvalues_opt are the same (using the same method as in the chunk above)
### -> YES, THEY ARE

deltaS_mg <- getDeltaS(xMg) %>% mutate(deltaSdiff = deltaS_fit - deltaS_sample) 
summary(deltaS_mg$deltaSdiff); sd(deltaS_mg$deltaSdiff)
deltaS_mg <- deltaS_mg %>% select(concept_a, concept_b, color1, color2, deltaS_sample, deltaS_fit) 
```

```{r}
# write_csv(deltaS_mg, "output/mg-deltaS.csv")
# write_csv(deltaS_us, "output/us-deltaS.csv")
```

Fitting alpha value from the US and MG data combined
```{r}
bothdf_agg <- rbind(select(mgdf, -conceptFr), usdf) %>%
  group_by(country, concept, color) %>%
  summarise(xbar = mean(rating), sd_sample = sd(rating), .groups = "drop")
  
a_both <- round(coef(getFitSd(bothdf_agg)), digits = 2) # a = 1.47
# Plotting the data points, and fitted curve
plotFitSd(a_both, bothdf_agg, "Both data combined")

sd_diff_both <- a_both*bothdf_agg$xbar*(1-bothdf_agg$xbar) - bothdf_agg$sd_sample; var(sd_diff_both)

i=9; a_both*bothdf_agg$xbar[i]*(1-bothdf_agg$xbar[i]) - bothdf_agg$sd_sample[i];
```


EVALUATING DELTA S deviation.
i.e comparing delatS values when using alpha_mg vs alpha_both

```{r eval- mg data}
# with alpha_mg
deltaS_mg_a_mg <- deltaS_mg %>% select(concept_a, concept_b, color1, color2, deltaS_sample, deltaS_fit) 
# diff(deltaS_mg_a_mg , deltaS_sample) is minimal

#  deltaS w/ alpha_both
deltaS_mg_a_both <- getXvalues_opt(mgdf_agg, concept_pairs, color_pairs, a_both) %>% 
  mutate(deltaX = (xbar1 + xbar4) - (xbar2 + xbar3)) %>% 
 getDeltaS() %>% mutate(deltaSdiff = deltaS_fit - deltaS_sample) 
summary(deltaS_mg_a_both$deltaSdiff); sd(deltaS_mg_a_both$deltaSdiff)
deltaS_mg_a_both <- deltaS_mg_a_both %>% select(concept_a, concept_b, color1, color2, deltaS_sample, deltaS_fit) 
# diff(deltaS_mg_a_both , deltaS_sample) is minimal

# delta w/ alpha_mg vs alpha_both
deltaS_mg_compare <- left_join(deltaS_mg_a_mg, deltaS_mg_a_both, by = c("concept_a", "concept_b", "color1", "color2")) %>% select(-c(deltaS_sample.x, deltaS_sample.y)) %>% 
  rename(deltaS_fit_a_mg = deltaS_fit.x, deltaS_fit_a_both = deltaS_fit.y) %>% 
  mutate(delatS_fit_a_diff = deltaS_fit_a_mg - deltaS_fit_a_both)
summary(deltaS_mg_compare$delatS_fit_a_diff); sd(deltaS_mg_compare$delatS_fit_a_diff)
# it looks like the deltaS diff between using a_mg and a_both is also minimal for MG datasets

# Also verigied if deltaX are the same (just for sanity check) and yes, they are
# main diff between mg_landscape and deltaS_mg datasets are in the canonical order of concepts and color pairs.
# but deltaX match regardless
```

## Cross-cultural analysis

```{r cross-cultural analysis}
# Finding four pairs of colors for each pair of concepts such that
## working with deltaS_mg and deltaS_us

dsmg <- deltaS_mg %>%
  select(-deltaS_sample) %>%
  rename(deltaSmg = deltaS_fit)
dsus <- deltaS_us %>%
  select(-deltaS_sample) %>%
  rename(deltaSus = deltaS_fit)

getPairs <- function(dsL, dsH) {
  dsmgL <- filter(dsmg, deltaSmg <= dsL) %>% select(-deltaSmg)
  dsmgH <- filter(dsmg, deltaSmg >= dsH) %>% select(-deltaSmg)
  dsusL <- filter(dsus, deltaSus <= dsL) %>% select(-deltaSus)
  dsusH <- filter(dsus, deltaSus >= dsH) %>% select(-deltaSus)

  mgL_usH <- inner_join(dsmgL, dsusH, by = c("concept1", "concept2", "color1", "color2")) %>% # common rows between the two dataframes
    rename(mgL_usH1 = color1, mgL_usH2 = color2)
  mgH_usH <- inner_join(dsmgH, dsusH, by = c("concept1", "concept2", "color1", "color2")) %>%
    rename(mgH_usH1 = color1, mgH_usH2 = color2)
  mgL_usL <- inner_join(dsmgL, dsusL, by = c("concept1", "concept2", "color1", "color2")) %>%
    rename(mgL_usL1 = color1, mgL_usL2 = color2)
  mgH_usL <- inner_join(dsmgH, dsusL, by = c("concept1", "concept2", "color1", "color2")) %>%
    rename(mgH_usL1 = color1, mgH_usL2 = color2)

  final <- mgL_usH %>%
    full_join(mgH_usH, by = c("concept1", "concept2"), relationship = "many-to-many") %>%
    full_join(mgL_usL, by = c("concept1", "concept2"), relationship = "many-to-many") %>%
    full_join(mgH_usL, by = c("concept1", "concept2"), relationship = "many-to-many") #  %>% drop_na()

  return(final)
}

# Lower and Higher thresholds for deltaS
dsL <- 0.1
dsH <- 0.7
final <- getPairs(dsL, dsH) %>% drop_na()
unique(final$concept1)
unique(final$concept2)
# find which columns of the dataframe only have NAs
names(final)[sapply(final, function(x) sum(is.na(x)) == length(x))]
final <- final %>% filter(!is.na(mgH_usL1), !is.na(mgH_usL2))

# unique combinations of two columns concept1 and concept2
final %>% distinct(concept1, concept2)
final %>%
  group_by(concept1, concept2) %>%
  summarize(freq = n())

## Adding color code to Final dataset
bcp37 <- read_csv("utils/color_values_bcp37.csv") # %>% select('name', 'color', 'hex')
```


```{r cross-cultural analysis}
write_csv(final, paste("output/", gsub("\\.", "", dsL), "-", gsub("\\.", "", dsH), "-final.csv", sep = ""))
```


```{r generate data for d3 visualization}
library(jsonlite)

# Sample data (to be replaced with actual data)
# df <- data.frame(
#   concept1 = c("banana", "banana", "test1"),
#   concept2 = c("love", "love", "test2"),
#   mgL_usH1 = c("WH", "SR", "LH1"),
#   mgL_usH2 = c("XR", "MR", "LH2"),
#   mgH_usH1 = c("MG", "WH", "HH1"),
#   mgH_usH2 = c("MR", "MG", "HH2"),
#   mgL_usL1 = c("SR", "MR", "MR"),
#   mgL_usL2 = c("MG", "WH", "MR"),
#   mgH_usL1 = c("WH", "SR", "SR"),
#   mgH_usL2 = c("MR", "MG", "XR"),
#   stringsAsFactors = FALSE
# )

## !!! NOTE The FILTER
df <- final # %>% filter(mgL_usH1 == "WH", mgL_usH2 == "DP", mgH_usH1=="DG", mgH_usH2 == "BK")

json_list <- list()
for (i in 1:nrow(df)) {
  # Extract values for the current row
  row_data <- df[i, ]

  # Define the JSON structure for the current row
  json_data <- list(
    concept1 = row_data$concept1,
    concept2 = row_data$concept2,
    mgL_usH = list(row_data$mgL_usH1, row_data$mgL_usH2),
    mgH_usH = list(row_data$mgH_usH1, row_data$mgH_usH2),
    mgL_usL = list(row_data$mgL_usL1, row_data$mgL_usL2),
    mgH_usL = list(row_data$mgH_usL1, row_data$mgH_usL2)
  )
  json_list[[i]] <- json_data
}

# Convert the entire list to JSON and write it to a single file
json_string <- toJSON(json_list, pretty = TRUE, auto_unbox = TRUE)
# print(json_string)

# write(json_string, file = "output/final-12-2026.json")
write(json_string, file = paste("output/", gsub("\\.", "", dsL), "-", gsub("\\.", "", dsH), "-final.json", sep = ""))

# df2 <- data.frame(
#   color = c("WH", "SR", "MR", "MG", "XR", "LH1", "LH2", "HH1", "HH2"),
#   hex = c("white", "red", "orange", "green", "purple", "lightyellow", "light", "HH1", "HH2"),
#   L = c(100, 70, 40, 30, 0, 0,0,0,0),
#   a = c(0, 70, -70, 0, 0,0,0,0,0),
#   b = c(0, 0, 70, -70, 0,0,0,0,0),
#   stringsAsFactors = FALSE
# )

df2 <- bcp37
# Function to create JSON-like entry for each color (more info about the color)
create_json_entry <- function(code) {
  entry <- df2 %>% filter(color == code)
  if (nrow(entry) > 0) {
    return(list(color = code, hex = entry$hex, L = entry$L, a = entry$a, b = entry$b))
    # return(list(color = code, hex = entry$hex, L = entry$L))
  } else {
    return(NULL) # Return NULL if the color code doesn't exist in df2
  }
}

# Replace color codes with JSON entries
df_replaced <- df %>%
  mutate(across(starts_with("mg"), ~ map(.x, create_json_entry)))

# Convert the transformed dataframe back to JSON
output_json <- toJSON(df_replaced, pretty = TRUE, auto_unbox = TRUE)

# cat(output_json)

df_combined <- df_replaced %>%
  mutate(mgL_usH = map2(mgL_usH1, mgL_usH2, ~ list(.x, .y))) %>%
  mutate(mgL_usL = map2(mgL_usL1, mgL_usL2, ~ list(.x, .y))) %>%
  mutate(mgH_usH = map2(mgH_usH1, mgH_usH2, ~ list(.x, .y))) %>%
  mutate(mgH_usL = map2(mgH_usL1, mgH_usL2, ~ list(.x, .y))) %>%
  select(-mgL_usH1, -mgL_usH2, -mgL_usL1, -mgL_usL2, -mgH_usH1, -mgH_usH2, -mgH_usL1, -mgH_usL2) # Remove the original keys

# Convert the modified dataframe back to JSON
output_json <- toJSON(df_combined, pretty = TRUE, auto_unbox = TRUE)

# Print the resulting JSON
# cat(output_json)
# write(output_json, file = "output/full-12-2026.json")
write(output_json, file = paste("output/", gsub("\\.", "", dsL), "-", gsub("\\.", "", dsH), "-full.json", sep = ""))
```
